Spatial variation in allometric growth of invasive lionfish has management implications

### 2019-01-30 docs/Manuscript.tex
Added all comments from reviewers, must proof read and send to Sean

\author[1]{Juan Carlos Villaseñor-Derbez}
\author[1]{Sean Fitzgerald}
\affil[1]{Bren School of Environmental Sciences and Management, University of California
  Santa Barbara, Santa Barbara, California, USA}
\corrauthor[1]{Juan Carlos Villaseñor-Derbez}{jvillasenor@ucsb.edu}

@@

\begin{abstract}
Lionfish (\textit{Pterois volitans / miles}) are an invasive species in
the Western Atlantic and the Caribbean. Improving management of invasive
lionfish populations requires accurate total biomass estimates, which
depend on accurate estimates of allometric growth. Sedentary species
like lionfish often exhibit high levels of spatial variation in life
history characteristics. We review 17 published length-weight
relationships for lionfish taken throughout their invasive range and
found substantial regional differences in allometric growth parameters.
The spatial pattern we observed is consistent with findings from other
studies focusing on genetics or age-at-length. We show that the use of
\textit{ex situ} parameters can result in up to a threefold under- or
overestimation of total weight, but using parameters from nearby regions
reduces this error. These findings can have major implications for
management in terms of predicting effects on local ecosystems,
evaluating the effectiveness of removal programs, or estimating biomass
available for harvest.
\end{abstract}

@@

Lionfish (\emph{Pterois volitans/miles} complex) are an invasive species
in the western Atlantic and Caribbean Sea, likely introduced through
liberation of aquarium-kept organisms \citep{betancurr_2011}. They are
the first invasive marine vertebrates established along the North
Atlantic Caribbean coasts
\citep{schofield_2009,schofield_2010,sabidoitza_2016} and their presence
has been labeled as a major marine invasion because they threaten local
biodiversity, spread rapidly, and are difficult to manage
\citep{hixon_2016}. Lionfish have established invasive populations in
coral reefs, estuaries, mangroves, hard-bottomed areas, and mesophotic
reefs
\citep{barbour_2010,jud_2011,muoz_2011,claydon_2012,andradibrown_2017,gress_2017}.

A substantial amount of research describes lionfish impacts throughout
its invaded range. A meta-analysis by \citet{peake_2018} showed that
invasive lionfish prey on at least 167 different species across the
tropical and temperate North Atlantic. Their feeding behavior and high
consumption rates can reduce recruitment and population sizes of native
reef-fish species, and can further endanger reef fish
(\citet{green_2012,rocha_2015}; but see \citet{hackerott_2017}). For
example, field experiments by \citet{albins_2008} showed that lionfish
establishment led to reduced recruitment of native fishes by nearly 80\%
over a five-week period in Florida. \citet{green_2012} reported that
prey fish biomass declined by 65\% over two years as lionfish biomass
increased along Bahamian coral reefs. Their trophic impacts can be
minimized if local lionfish biomass is controlled by culling
\citep{ariasgonzalez_2011}.

Governments and non-profit organizations have sought to reduce lionfish
densities through removal programs and incentivizing its consumption
\citep{chin_2016}. In some cases, these have shown to significantly
reduce --but not quite eliminate-- lionfish abundances at local scales
\citep{deleon_2013,sandel_2015}. Complete eradication of lionfish
through fishing is unlikely because of their rapid recovery rates and
ongoing recruitment to shallow-water areas from persistent populations
in mesophotic ecosystems \citep{barbour_2011,andradibrown_2017}.
However, promoting lionfish consumption might create a level of demand
capable of incentivizing a stable fishery while controlling
shallow-water populations, thus creating alternative livelihoods and
avoiding further impacts to local biota.

@@

Outcomes of previous studies suggest lionfish are likely to exhibit
spatial heterogeneity in the length-weight relationship, which we
summarize in two main causes. First, culling programs are effective in
reducing local adult populations largely because lionfish exhibit high
levels of site fidelity and small home ranges
\citep{Fishelson_1997,kochzius_2005,jud_2012,cote_2014}. It is know that
fish with sedentary behavior are likely to exhibit high levels of
spatial variation in important life history characteristics such as
growth or natural mortality rates
\citep{gunderson_2008,hutchinson_2008,wilson_2012,guan_2013}. Second,
genetic analysis of lionfish suggests biological differences due to the
existence of two genetically distinct invasive subpopulations between
the northwest Atlantic and the Caribbean \citep{betancurr_2011}.
Site-specific parameters are necessary to accurately estimate biomass
when allometric relationships are spatially variable, and this
variability is increasingly important when estimating the potential
effectiveness of lionfish culling programs
\citep{barbour_2011,morris_2011,cote_2014,johnston_2015}. However, the
region-wide differences in allometric growth parameters has remained
unexplored for lionfish, despite the large number of site-specific
studies reporting the length-weight relationship.

Here, we compare previously published length-weight relationships for
lionfish populations in North Carolina, Northern and Southern Gulf of
Mexico, the Southern Mexican Caribbean, Bahamas, Little Cayman, Jamaica,
Bonaire, Puerto Rico, and Costa Rica
\citep{barbour_2011,darling_2011,deleon_2013,fogg_2013,dahl_2014,edwards_2014,toledohernndez_2014,sandel_2015,aguilarperera_2016,sabidoitza_2016,sabidoitz_2016,chin_2016}.
We also collected lionfish length and weight data in the central Mexican
Caribbean and report the first allometric growth equation for this
region. The objective of this paper is to describe the spatial pattern
of length-weight relationships of lionfish across the Caribbean and
Western Atlantic and to discuss implications of these spatial
differences.

@@

We reviewed 12 published studies and obtained 17 length-weight
relationships for the North Atlantic (n = 1), Gulf of Mexico (n = 7),
and Caribbean (n = 9, Table \ref{tab:all_params}, Fig
\ref{fig:all_allo}). We collected information on sampling methods, sex
differentiation, location, and depth ranges from each study when
available. Only two studies reported parameters for each gender
\citep{aguilarperera_2016,fogg_2013}, so we assumed both genders were
included in a study if gender was unspecified. Reviewed studies
presented information for organisms obtained at depths between 0.5 m and
57 m. Three studies explicitly stated that their organisms were sampled
with pole spears
\citep{dahl_2014,aguilarperera_2016,chin_2016,sabidoitz_2016}, and five
studies mentioned that some of their organisms were obtained with pole
spears (or other type of harpoon) but also hand-held nets or fish traps
\citep{barbour_2011,fogg_2013,edwards_2014,toledohernndez_2014,sandel_2015,sabidoitza_2016,sabidoitz_2016},
and two studies did not specify how organisms were sampled
\citep{darling_2011,deleon_2013}. \citet{fogg_2013} use spineless weight
in their calculations, so their parameters likely underestimated total
weight. Since no spineless to total weight conversions were available,
these parameters were taken as reported.

@@

We also used data from \citet{villaseorderbez_2014}, who collected
organisms from 10 sampling sites along the central Mexican Caribbean
coast in 2010 (Supplementary Table 1). Sampling locations included wall
and carpet reefs at depths between 5.7 m and 38.1 m. All observed
lionfish (n = 109) were collected using hand nets and numbered
collection bottles. The use of hand nets prevented any weight loss due
to bleeding and allowed better representation of small sizes by
eliminating gear selectivity. Organisms were euthanized via pithing and
Total Length (TL; mm) and Total Weight (TW; g) were recorded.

@@

Transforming this equation via base-10 logarithms we obtain:

\begin{equation}
\label{eq:log-alo}
log_{10}(TW) = b\times log_{10}(TL) + log_{10}(a)
\end{equation}

This can be simplified and re-written as:

\begin{equation}
\label{eq:log-alo-trans}
Y = bX + c
\end{equation}

Where \(Y = log_{10}(TW)\), \(X = log_{10}(TL)\), and
\(c = log_{10}(a)\). The coefficients (\(c\) and \(b\)) were estimated
with an Ordinary Least Squares Regression and heteroskedastic-robust
standard error correction \citep{zeileis_2004}. When the \(b = 3\), it
is said that the organism exhibits a perfect isometric growth, so the
\(b\) coefficient was tested against the null hypothesis of isometric
growth (\emph{i.e.} \(H_0: b = 3\)). Coefficients were tested with a
two-tailed Student's t, and the significance of the regression was
corroborated with an F-test.

Some of the reviewed studies inconsistently defined \(a\) as either the
ponderal index from Eq. \ref{eq:allometric} or the y-intercept (\(c\))
from Eq. \ref{eq:log-alo-trans}. Other studies incorrectly reported
parameters as mm-to-g conversions when they were in fact cm-to-g
conversions. We standardized each study by converting coefficients and
report all parameters as TL(mm) to TW (gr) conversions. Locations where
allometric studies have been performed are shown in Figure \ref{fig:map}
and summarized in Table \ref{tab:all_params}.

We obtained a total of 18 parameter pairs by combining length-weight
parameters extracted from the literature and the additional pair
calculated here. We used the central Mexican Caribbean as a case study
of how the use of \emph{ex situ} parameters influences the accuracy of
weight estimates for lionfish. We estimated TW from the TL observations
we collected in the central Mexican Caribbean (n = 109, with
\(TL \in (34, 310)\)) using each of the 18 parameter pairs and divided
predicted weights by known observed weights to obtain a simple measure
of over- or underestimation. Difference in mean weight ratios across the
different parameter pairs were tested with a one-way analysis of
variance (ANOVA) and Tukey's test was used for \emph{post-hoc} tests.
All analyses were performed in R version 3.5.1 \citep{rcore_2018}. Raw
data and code used in this work are available on github.

@@

The length-weight relationship for organisms from the central Mexican
Caribbean resulted in the coefficient values
\(a = 3.2056297\times 10^{-6}\), \(b = 3.2347391\) and
\(c = -5.4940866\) (\(R^2 = 0.977\), F(df = 1; 107) = 6928.67,
\(p < 0.001\)). The allometric factor (\(b\)) was significantly
different from \(b = 3\) (\(t(107) = 6.04; p<0.001\)) indicating that
lionfish present allometric growth. The length-weight coefficients
estimated in this study were within the range identified by studies in
other regions (Table \ref{tab:all_params}). Figure \ref{fig:l-w-carib}
shows the relationship between TL and TW for this region, and model fit
statistics are presented in Table \ref{tab:reg_table}.

@@

\begin{table}[!htbp] \centering 
  \caption{\label{tab:reg_table}Coefficients of the linear model fit to Eq \ref{eq:log-alo-trans}. Numbers in parenthesees represent heteroskedastic-robust standard errors.} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{1}{c}{$log_{10}(TW)$} \\ 
\cline{2-2} 
\hline \\[-1.8ex] 
 c & $-$5.494 (0.083)$^{***}$ \\ 
  b & 3.235 (0.039)$^{***}$ \\ 
 \hline \\[-1.8ex] 
F Statistic & 6928.67*** (df = 1; 107) \\ 
Observations & 109 \\ 
Adjusted R$^{2}$ & 0.976 \\ 
Residual Std. Error & 0.096 (df = 107) \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

There were significant differences in our predicted weights for the
central Mexican Caribbean when using the different pairs of parameters
(\(F(df = 17; 1944) = 61.55; p < 0.001\)). The lowest weight estimates
resulted from using the allometric parameters from Banco Chinchorro in
the Caribbean, with mean \(\pm\) SD of 40.37 \(\pm\) 58.74 gr
\citep{sabidoitz_2016}, and the highest weight estimates came from the
Northern Atlantic with 73.76 \(\pm\) 96.11 gr \citep{barbour_2011}. To
put this in context, true observed weights were 52.56 \(\pm\) 76.58 gr.
These correspond to predicted-to-observed weights ratios of 0.80 \(\pm\)
0.19 and 1.76 \(\pm\) 0.50 (mean \(\pm\) SD), respectively.

The calculated ratio of predicted-to-observed weight ranged from 0.36 to
3.51, indicating that \emph{ex situ} parameters can result in major
under- and overestimation. Tukey's \emph{post-hoc} test suggests that
weight ratios for the central Mexican Caribbean were not different from
those obtained with parameters from Little Cayman, the Bahamas, and some
sites in the Gulf of Mexico (Tukey's HSD \(p > 0.05\)). Weight estimates
using parameters from the Gulf of Mexico and North-Western Atlantic were
higher on average than those from the Caribbean (Fig
\ref{fig:all_allo}). The average (\(\pm\) SD) predicted-to-observed
weight ratios from these three regions were 1.24 \(\pm\) 0.309, 1.76
\(\pm\) 0.496, and 1.17 \(\pm\) 0.398, respectively.
Predicted-to-observed weight ratios are presented in Figure
\ref{fig:bio_ratio}. Spineless weight parameters from \citet{fogg_2013}
still produced predicted-to-observed weight ratios \textgreater{} 1.

\begin{table}

\caption{\label{tab:unnamed-chunk-7}\label{tab:all_params}Summary of 18 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both genders combined (B), a = scaling parameter for Eq. 1 (presented in $\times 10^{-5}$), c = y-intercept for Eq. 3, b = exponent or slope for Eq. 1 or Eq. 3, respectively. The Fit column contains the reported $R^2$ of the model fit.}
\centering
\begin{tabular}[t]{llllrrll}
\toprule
Region & Sex & n & a & b & c & Fit & Reference\\
\midrule
Caribbean & B & 458 & 3.6 & 2.81 & -4.44 & - & Sandel et al., 2015\\
Caribbean & B & 419 & 2.8 & 2.85 & -4.56 & 0.8715 & Chin et al., 2016\\
Caribbean & B & 1450 & 2.3 & 2.89 & -4.64 & 0.96 & de Leon et al., 2013\\
Caribbean & B & 1887 & 0.3 & 3.24 & -5.52 & 0.97 & Edwards et al., 2014\\
Caribbean & B & - & 0.25 & 3.29 & -5.60 & - & Darling et al., 2011\\
\addlinespace
Caribbean & B & 2143 & 0.52 & 3.18 & -5.28 & 0.9907 & Sabido-Itza et al., 2016\\
Caribbean & B & 227 & 0.8 & 3.11 & -5.10 & 0.958 & Toledo-Hernández et al., 2014\\
Caribbean & B & 449 & 0.23 & 3.25 & -5.64 & 0.97 & Sabido-Itza et al., 2016b\\
Caribbean & B & 368 & 0.32 & 3.19 & -5.50 & 0.98 & Sabido-Itza et al., 2016b\\
Caribbean & B & 109 & 0.32 & 3.23 & -5.49 & 0.9766 & This study\\
\addlinespace
GoM & B & 934 & 0.21 & 3.34 & -5.68 & 0.98 & Dahl \& Patterson, 2014\\
GoM & B & 472 & 0.29 & 3.30 & -5.54 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & F & 67 & 0.12 & 3.47 & -5.93 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & M & 59 & 0.42 & 3.23 & -5.38 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & B & 582 & 0.14 & 3.43 & -5.86 & 0.99 & Fogg et al., 2013\\
\addlinespace
GoM & M & 119 & 0.27 & 3.31 & -5.57 & 0.97 & Fogg et al., 2013\\
GoM & F & 115 & 0.68 & 3.14 & -5.17 & 0.94 & Fogg et al., 2013\\
North Atlantic & B & 774 & 2.9 & 2.89 & -4.54 & - & Barbour et al., 2011\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/fit2-1.pdf}
\caption{\label{fig:all_allo}Length-weight relationships (n = 18) for 12
studies and this study. Colors indicate studies from which the
parameters were extracted. Dotted, dashed and solid lines show models
for males, females, and combined sexes, respectively. The dashed black
line represents the relationship estimated in this study.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/pred_obs-1.pdf}
\caption{\label{fig:bio_ratio}Violin plot of predicted-to-observed
weight ratios for 18 pairs of allometric parameters. Blue circles
indicate median values and Like letters indicate values that do not
differ significantly.}
\end{figure}

@@

Our results suggest that lionfish exhibit highly variable allometric
relationships across the invaded range, and that this variation is
related to space. Moreover, we shot that the use of \emph{ex situ}
parameters may lead to highly biased weight estimates. Our comparison of
observed weights to those predicted with locally-informed parameters and
\emph{ex situ} parameters showed that weight can be overestimated by
more than a three-fold, and highlights the need to use local
information. Here we discuss the implications of our findings and
highlight potential future research directions.

We detected substantial differences in weight-at-length between
organisms from the Caribbean, Gulf of Mexico, and North-Western
Atlantic. Groupings of predicted-to-observed weight ratios aligned with
the spatial distribution of the examined studies, suggesting that these
differences are mediated by space. These regional allometric differences
mirror similar patterns in age-at-length of lionfish across both their
invaded and native regions \citep{pusack_2016}. Variation may be driven
by genetics or by organisms' exposure to distinct environmental
conditions. For example, \citet{betancurr_2011} used mitochondrial DNA
to demonstrate the existence of two distinct population groups,
identified as the ``Caribbean group'' and ``Northern Group'', and
\citet{fogg_2015} alternatively suggested that age-at-length differences
may be climate-driven. Differences in weight-at-length could also
reflect differential energy input or usage, or a combination of both.
Future research is needed to determine which processes are at work here.

@@

The results presented here have major implications for management. For
example, \citet{edwards_2014} simulated a lionfish culling program under
two scenarios, one using length-at-age and length-to-weight parameters
from North Carolina and one using parameters from Little Cayman. Their
results show that using different parameters caused up to a four-year
difference in the time required for the simulated lionfish population to
recover to 90\% of its initial biomass after removals ceased. Here, we
show that using one set of length-weight parameters versus another for a
given length can result in more than a threefold under- or
overestimation of total weight. These spatially-driven differences
become especially important when allocating resources for lionfish
removal programs, incentivizing lionfish fisheries as a source of
alternative livelihoods, or estimating ecosystem impacts. Research
efforts focused on invasive lionfish populations need to use parameters
calculated for their region to the extent possible, or at least use
reasonable sets of different parameters that provide upper and lower
bounds in their results.

>>>>

\author[1]{Juan Carlos Villaseñor-Derbez}
\author[1]{Sean Fitzgerald}
\affil[1]{Bren School of Environmental Sciences and Management, University of California
  Santa Barbara, Santa Barbara, California, USA}
\corrauthor[1]{Juan Carlos Villaseñor-Derbez}{juancarlos@ucsb.edu}

@@

\begin{abstract}
Lionfish (\textit{Pterois volitans / miles}) are an invasive species in
the Western Atlantic and the Caribbean. Improving management of invasive
lionfish populations requires accurate total biomass estimates, which
depend on accurate estimates of allometric growth. Sedentary species
like lionfish often exhibit high levels of spatial variation in life
history characteristics. We review 17 published length-weight
relationships for lionfish taken throughout their invasive range and
found significant regional differences in allometric growth parameters.
The spatial pattern we observed is consistent with findings from other
studies focusing on genetics or length-at-age. We show that the use of
\textit{ex situ} parameters can result in up to a threefold under- or
overestimation of total weight, but using parameters from nearby regions
reduces this error. These findings can have substantial implications for
management in terms of predicting effects on local ecosystems,
evaluating the effectiveness of removal programs, or estimating biomass
available for harvest.
\end{abstract}

@@

Lionfish (\emph{Pterois volitans/miles} complex) are an invasive species
in the Western Atlantic Ocean and Caribbean Sea, likely introduced
through release of aquarium-kept organisms \citep{betancurr_2011}. They
are the first invasive marine vertebrates established along these coasts
\citep{schofield_2009,schofield_2010,sabidoitza_2016} and their presence
has been labeled as a major marine invasion because they threaten local
biodiversity, spread rapidly, and are difficult to manage
\citep{hixon_2016}. Lionfish have established invasive populations in
coral reefs, estuaries, mangroves, hard-bottomed areas, and mesophotic
reefs
\citep{barbour_2010,jud_2011,muoz_2011,claydon_2012,andradibrown_2017,gress_2017}.

A substantial amount of research describes lionfish impacts throughout
its invaded range. A meta-analysis by \citet{peake_2018} showed that
invasive lionfish prey on at least 167 different species across the
tropical and temperate Western Atlantic. Their feeding behavior and high
consumption rates can reduce recruitment and population sizes of native
reef-fish species, and can further endanger reef fish
\citep{green_2012,rocha_2015}; but see \citet{hackerott_2017}. For
example, field experiments by \citet{albins_2008} showed that lionfish
establishment led to reduced recruitment of native fishes by nearly 80\%
over a five-week period in the Bahamas. \citet{green_2012} reported that
prey fish biomass declined by 65\% over two years as lionfish biomass
increased along Bahamian coral reefs. However, their trophic impacts can
be minimized if local lionfish biomass is controlled by culling
\citep{ariasgonzalez_2011}.

Governments and non-profit organizations have sought to reduce lionfish
densities through removal programs and incentivizing its consumption
\citep{chin_2016}. In some cases, these have shown to significantly
reduce --but not quite eliminate-- lionfish abundances at local scales
\citep{deleon_2013,sandel_2015}. Complete eradication of lionfish
through fishing is unlikely because of their rapid recovery rates and
ongoing recruitment to shallow-water areas from persistent populations
in mesophotic ecosystems \citep{barbour_2011,andradibrown_2017}.
However, promoting lionfish consumption might create a level of demand
capable of incentivizing a stable fishery while controlling
shallow-water populations, thus creating alternative livelihoods and
avoiding further negative effects to local biota.

@@

Outcomes of previous studies suggest lionfish are likely to exhibit
spatial heterogeneity in the length-weight relationship, which we
summarize in two main causes. First, culling programs are effective in
reducing local adult populations largely because lionfish exhibit high
levels of site fidelity and small home ranges
\citep{Fishelson_1997,kochzius_2005,jud_2012,cote_2014}. It is known
that fish with sedentary behavior are likely to exhibit high levels of
spatial variation in important life history characteristics such as
growth or natural mortality rates
\citep{gunderson_2008,hutchinson_2008,wilson_2012,guan_2013}. Second,
genetic analysis of lionfish suggests biological differences due to the
existence of two genetically distinct invasive subpopulations between
the Western Atlantic and the Caribbean \citep{betancurr_2011}. The large
number of site-specific studies reporting the length-weight relationship
of lionfish provide variable estimates. These differences may be
increasingly important when estimating the potential effectiveness of
lionfish culling programs
\citep{barbour_2011,morris_2011,cote_2014,johnston_2015}. However, the
extent to which using \emph{ex situ} parameters to estimate total weight
from length observations remains unexplored.

Here, we use previously published length-weight relationships for
lionfish populations in North Carolina, Northern and Southern Gulf of
Mexico, the Southern Mexican Caribbean, Bahamas, Little Cayman, Jamaica,
Bonaire, Puerto Rico, and Costa Rica to quantify the magnitude of the
error caused by using \emph{ex situ} parameters to estimate lionfish
weight from length observations. We also collected lionfish length and
weight data in the central Mexican Caribbean and report the first
length-weight relationship for this region.

@@

We reviewed 12 published studies and obtained 17 length-weight
relationships for the Western Atlantic (n = 2), Gulf of Mexico (n = 7),
and Caribbean (n = 8, Table \ref{tab:all_params}, Fig \ref{fig:map}). We
collected information on sampling methods, sex differentiation,
location, and depth ranges from each study when available. Only two
studies reported parameters for each gender
\citep{aguilarperera_2016,fogg_2013}, so we assumed both genders were
included in a study if gender was unspecified. Reviewed studies
presented information for organisms obtained at depths between 0.5 m and
57 m. Four studies explicitly stated that their organisms were sampled
with pole spears
\citep{dahl_2014,aguilarperera_2016,chin_2016,sabidoitz_2016}, and six
studies mentioned that some of their organisms were obtained with pole
spears (or other type of harpoon) but also hand-held nets or fish traps
\citep{barbour_2011,fogg_2013,edwards_2014,toledohernndez_2014,sandel_2015,sabidoitza_2016},
and two studies did not specify how organisms were sampled
\citep{darling_2011,deleon_2013}.

@@

We also used data from \citet{villaseorderbez_2014}, who collected
organisms from 10 sampling sites along the central Mexican Caribbean
coast in the Summer of 2010 (Supplementary Table 1). Sampling locations
included wall and carpet reefs at depths between 5.7 m and 38.1 m. All
observed lionfish (n = 109) were collected using hand nets and numbered
collection bottles. The use of hand nets prevented any weight loss due
to bleeding and allowed better representation of small sizes by avoiding
gear selectivity. Organisms were euthanized via pithing and Total Length
(TL; mm) and Total Weight (TW; g) were recorded.

@@

The above equation was linearized using \(log_{10}\)-transformation. The
coefficients were estimated with an Ordinary Least Squares Regression,
and heteroskedastic-robust standard error correction was applied
\citep{zeileis_2004}. When \(b = 3\), it is said that the organism
exhibits a perfect isometric growth, so the \(b\) coefficient was tested
against the null hypothesis of isometric growth (\emph{i.e.}
\(H_0: b = 3\)). Coefficients were tested with a two-tailed Student's t,
and the significance of the regression was corroborated with an F-test.

Some of the reviewed studies inconsistently defined \(a\) as either the
ponderal index from Eq. \ref{eq:allometric} or the y-intercept from the
linearized log-transformed equation. Other studies incorrectly reported
parameters as mm-to-g conversions when they were in fact cm-to-g
conversions. We standardized each study by converting coefficients and
report all parameters as TL(mm) to TW (g) conversions. Locations where
allometric studies have been performed are shown in Figure \ref{fig:map}
and summarized in Table \ref{tab:all_params}.

We obtained a total of 18 parameter pairs by combining length-weight
parameters extracted from the literature and the additional pair
calculated here. Recall that the objective of this study is not to
describe population-level variations, but rather to estimate how the use
of \emph{ex situ} parameters influences weight estimates. We used the
central Mexican Caribbean as a case study of how the use of \emph{ex
situ} parameters influences the accuracy of weight estimates for
lionfish. We estimated \(TW\) from the \(TL\) observations we collected
in the central Mexican Caribbean (n = 109, with \(TL\) ranging from 34
mm to 310 mm) using each of the 18 parameter pairs and divided predicted
weights by known observed weights to obtain a simple measure of over- or
underestimation. Difference in mean weight ratios were tested with an
analysis of covariance (ANCOVA) using study and \(TL\) as covariates.
Ratios were logit-transformed prior to analysis, and a \emph{post-hoc}
Tukey's test was used to identify groups where mean ratios did not
differ. All analyses were performed in R version 3.5.2
\citep{rcore_2018}. Raw data and code used in this work are available on
github at github.com/jcvdav/lionfish\_biometry.

@@

The length-weight relationship for organisms from the central Mexican
Caribbean resulted in the coefficient values
\(a = 3.2056297\times 10^{-6}\), \(b = 3.2347391\) and
\(c = -5.4940866\) (\(R^2 = 0.977\), F(df = 1; 107) = 6928.67,
\(p < 0.001\)). The allometric factor (\(b\)) was significantly
different from \(b = 3\) (\(t(107) = 6.04; p<0.001\)) indicating that
lionfish present allometric growth. The length-weight coefficients
estimated in this study were within the range identified by studies in
other regions (Table \ref{tab:all_params}). Figure \ref{fig:l-w-carib}
shows the relationship between TL and TW for this region.

@@

There were significant differences in our predicted weights for the
central Mexican Caribbean when using each of the different pairs of
parameters (\(F(df = 17; 1944) = 61.55; p < 0.001\)). The lowest weight
estimates for the observed lengths resulted from using the allometric
parameters from Banco Chinchorro in the Caribbean, with mean \(\pm\) SD
of 40.37 \(\pm\) 58.74 g \citep{sabidoitz_2016}. In contrast, the
highest weight estimates came from the Western Atlantic with 73.76
\(\pm\) 96.11 g \citep{barbour_2011}. To put this in context, true
observed weights have a mean of 52.56 \(\pm\) 76.58 g. Weigths predicted
from these extreme parameters correspond to mean \(\pm\) SD
predicted-to-observed weights ratios of 0.80 \(\pm\) 0.19 and 1.76
\(\pm\) 0.50 (mean \(\pm\) SD), respectively. At its extremes, other
parameters resulted in under- and overestimations in the order of 0.36
to 3.51 of the actual observed weight, indicating that \emph{ex situ}
parameters can result in substantial weight under- and overestimation.

Tukey's \emph{post-hoc} test suggests that weight ratios for the central
Mexican Caribbean were not different from those obtained with parameters
from Little Cayman, the Bahamas, and some sites in the Gulf of Mexico
(Tukey's HSD \(p > 0.05\)). Weight estimates using parameters from the
Gulf of Mexico and Western Atlantic were higher on average than those
from the Caribbean (Fig \ref{fig:all_allo}). The average (\(\pm\) SD)
predicted-to-observed weight ratios from these three regions were 1.24
\(\pm\) 0.309, 1.76 \(\pm\) 0.496, and 1.17 \(\pm\) 0.398, respectively.
Predicted-to-observed weight ratios are presented in Figure
\ref{fig:bio_ratio}.

\begin{table}[t]

\caption{\label{tab:unnamed-chunk-7}\label{tab:all_params}Summary of 18 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both genders combined (B), a = scaling parameter (presented in $\times 10^{-5}$), b = exponent.}
\centering
\begin{tabular}{llllrll}
\toprule
Region & Sex & n & a & b & \$R\textasciicircum{}2\$ & Reference\\
\midrule
Western Atlantic & B & 774 & 2.9 & 2.89 & - & Barbour et al., 2011\\
Western Atlantic & B & - & 0.25 & 3.29 & - & Darling et al., 2011\\
GoM & B & 934 & 0.21 & 3.34 & 0.98 & Dahl \& Patterson, 2014\\
GoM & B & 472 & 0.29 & 3.30 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & F & 67 & 0.12 & 3.47 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
\addlinespace
GoM & M & 59 & 0.42 & 3.23 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & B & 582 & 0.14 & 3.43 & 0.99 & Fogg et al., 2013\\
GoM & M & 119 & 0.27 & 3.31 & 0.97 & Fogg et al., 2013\\
GoM & F & 115 & 0.68 & 3.14 & 0.94 & Fogg et al., 2013\\
Caribbean & B & 458 & 3.6 & 2.81 & - & Sandel et al., 2015\\
\addlinespace
Caribbean & B & 419 & 2.8 & 2.85 & 0.87 & Chin et al., 2016\\
Caribbean & B & 1450 & 2.3 & 2.89 & 0.92 & de Leon et al., 2013\\
Caribbean & B & 1887 & 0.3 & 3.24 & 0.97 & Edwards et al., 2014\\
Caribbean & B & 2143 & 0.52 & 3.18 & 0.99 & Sabido-Itza et al., 2016\\
Caribbean & B & 227 & 0.8 & 3.11 & 0.96 & Toledo-Hernández et al., 2014\\
\addlinespace
Caribbean & B & 449 & 0.23 & 3.25 & 0.97 & Sabido-Itza et al., 2016b\\
Caribbean & B & 368 & 0.32 & 3.19 & 0.98 & Sabido-Itza et al., 2016b\\
Caribbean & B & 109 & 0.32 & 3.23 & 0.98 & This study\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/fit2-1.pdf}
\caption{\label{fig:all_allo}log-log transformation of the length-weight
relationships (n = 18) for 12 studies and this study. The curves are
shown for the range of lengths reported in each study (See Supplementary
Table 2); when ranges were not present, we use the ones found in this
study (34mm - 310 mm). Colors indicate studies from which the parameters
were extracted. Dotted, dashed and solid lines show models for males,
females, and combined sexes, respectively. The dashed black line
represents the relationship estimated in this study. There are two solid
green lines for Sabido-Itza et al, 2016b, one for each of the two sites
for which they report parameters. A log-log version of this figure is
presented in Figure S4.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/pred_obs-1.pdf}
\caption{\label{fig:bio_ratio}Violin plot of predicted-to-observed
weight ratios for 18 pairs of allometric parameters. Blue circles
indicate median values and like letters indicate values that do not
differ significantly. For Sabido-Itza et al, 2016b, BC and PNAX make
reference to Banco Chinchorro and Parque Nacional Arrecifes de Xcalak,
two sites for which they report parameters.}
\end{figure}

@@

Our results suggest that lionfish exhibit highly variable allometric
relationships across the invaded range, and that this variation is
spatially heterogeneous and relevant for management of the invasion.
Moreover, we show that the use of \emph{ex situ} parameters may lead to
highly biased weight estimates. Our comparison of observed weights to
those predicted with locally-informed parameters and \emph{ex situ}
parameters showed that weight can be overestimated by more than a
three-fold, and highlights the need to use local information. Here we
discuss the implications of our findings and highlight potential future
research directions.

@@

We detected substantial differences in weight-at-length between
organisms from the Caribbean, Gulf of Mexico, and Western Atlantic.
Groupings of predicted-to-observed weight ratios identified in our
\emph{post hoc} testing aligned with the spatial distribution of the
examined studies, suggesting that these differences may be mediated by
space. These regional allometric differences mirror similar patterns in
age-at-length of lionfish across both their invaded and native regions
\citep{pusack_2016}. Variation may be driven by genetics or by
organisms' exposure to distinct environmental conditions. For example,
\citet{betancurr_2011} used mitochondrial DNA to demonstrate the
existence of two distinct population groups, identified as the
``Caribbean group'' and ``Northern Group'', and \citet{fogg_2015}
alternatively suggested that age-at-length differences may be driven by
the environment.

We might be inclined to attribute all variation to the spatial origin of
these parameters. However, these were not only collected for different
locations, but also using a range of different sampling methods and for
different points in time (See Supplementary Table 2 for an extended
version of Table 1). While we are not able to evaluate how these factors
influence previous estimates (raw data from all studies would be
needed), it is certain that the lack of locally-calculated parameters
may induce significant bias when calculating weight from length
observations. Differences in weight-at-length could also reflect
differential energy input or usage, or a combination of both. The
magnitude of the bias and our lack of understanding of the source of
variation hihglights the need to simultaneously collect length-weight
information across the invaded range to test for spatially-induced
patterns and link these to previously suggested environmental and
genetic structures. Such an endeavor would provide insights into
lionfish biology and better inform management.

Applying parameters estimates to lengths that may be outside the range
of lengths originally used to estimate the parameters may also induce
variation lead to over- and underestimation. Our smallest observed
length was for an organism 34 mm in TL. However, only two studies
estimated parametrs with smaller organisms
\citep{sabidoitza_2016,edwards_2014}. On the other hand, our largest
organism had a TL of 310 mm, which is well within the range of all other
studies (maximum observed lengths varied from 325 mm to 475 mm; See
Supplementary Table 2). Due to the exponential nature of the allometric
relationship, the error is higher when extrapolation is done for TLs
that are larger than the maximum. This means that not only must managers
use locally-informed data, but that these local data must also include
the full range of lengths present in the region to reduce error caused
by extrapolation.
